Implementation and verification of a data structure using dependent types

### 2022-04-26 paper/main.tex
Add a detailed drawing of a binary leaf tree

\documentclass{article}
\usepackage[a4paper, total={19cm, 26cm}]{geometry}
\usepackage[pdfusetitle]{hyperref}
\usepackage{url}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage[backend=biber, style=alphabetic, sorting=none]{biblatex}
\addbibresource{main.bib}
\nocite{*}
\usepackage[usenames, dvipsnames, svgnames, x11names]{xcolor}
\usepackage{graphics}
\usepackage{scalerel}
\usepackage{accsupp}
\usepackage{stmaryrd}
\usepackage{listings}
\usepackage{lstcoq}
\usepackage{tikz}
\usetikzlibrary{matrix}
\usetikzlibrary{arrows.meta}
\tikzstyle{node_matrix} = [matrix of nodes, nodes in empty cells, nodes = {anchor = center, minimum size = 20, draw, fill = blue!10}, anchor = west]
\tikzstyle{arrow} = [[-{>[length=7, width=5]}]
\tikzstyle{leaf_tree} = [level distance = 25, level 1/.style = {sibling distance = 40}, level 2/.style = {sibling distance = 20}]
\tikzstyle{leaf_tree_leaf} = [minimum size = 0, circle, draw]
\tikzstyle{leaf_tree_inner_node} = [minimum size = 0, circle, draw, inner sep = 1, fill = black]
\newcommand{\bigvarnothing}[0]{%
    \scalebox{3}{%
        \begin{tikzpicture}%
            \clip (-0.114, -0.1) rectangle (0.114, 0.128);
            \node [style = {inner sep = 0}, text = black!35] {$\varnothing$};
        \end{tikzpicture}%
    }%
}
\newcommand{\completebinaryleaftreedepthzero}[1]{%
    \begin{tikzpicture}%
        \begin{scope} [leaf_tree]
            \node [style = {leaf_tree_leaf}] {#1};
        \end{scope}
    \end{tikzpicture}%
}
\newcommand{\completebinaryleaftreedepthone}[2]{%
    \begin{tikzpicture}%
        \begin{scope} [leaf_tree]
            \node [style = {leaf_tree_inner_node}] {}
                child {
                    node [style = {leaf_tree_leaf}] {#1}
                }
                child {
                    node [style = {leaf_tree_leaf}] {#2}
                };
        \end{scope}
    \end{tikzpicture}%
}
\newcommand{\completebinaryleaftreedepthtwo}[4]{%
    \begin{tikzpicture}%
        \begin{scope} [leaf_tree]
            \node [style = {leaf_tree_inner_node}] {}
                child {
                    node [style = {leaf_tree_inner_node}] {}
                    child {
                        node [style = {leaf_tree_leaf}] {#1}
                    }
                    child {
                        node [style = {leaf_tree_leaf}] {#2}
                    }
                }
                child {
                    node [style = {leaf_tree_inner_node}] {}
                    child {
                        node [style = {leaf_tree_leaf}] {#3}
                    }
                    child {
                        node [style = {leaf_tree_leaf}] {#4}
                    }
                };
        \end{scope}
    \end{tikzpicture}%
}
\newenvironment{no_space_center}
    {\parskip=0pt\par\nopagebreak\centering}
    {\par\noindent\ignorespacesafterend}
\newenvironment{no_padding_center}
    {\parskip=0pt\par\nopagebreak\centering}
    {\par\noindent}
\newcommand*{\llbrace}{%
    \BeginAccSupp{method=hex,unicode,ActualText=2983}%
    \textnormal{\usefont{OMS}{lmr}{m}{n}\char102}%
    \mathchoice{\mkern-4.05mu}{\mkern-4.05mu}{\mkern-4.3mu}{\mkern-4.8mu}%
    \textnormal{\usefont{OMS}{lmr}{m}{n}\char106}%
    \EndAccSupp{}%
}
\newcommand*{\rrbrace}{%
    \BeginAccSupp{method=hex,unicode,ActualText=2984}%
    \textnormal{\usefont{OMS}{lmr}{m}{n}\char106}%
    \mathchoice{\mkern-4.05mu}{\mkern-4.05mu}{\mkern-4.3mu}{\mkern-4.8mu}%
    \textnormal{\usefont{OMS}{lmr}{m}{n}\char103}%
    \EndAccSupp{}%
}
\newsavebox{\lXbrace}
\savebox{\lXbrace}{$\llbrace$}
\newsavebox{\rXbrace}
\savebox{\rXbrace}{$\rrbrace$}
\def\lxbrace{\scalerel*{\usebox{\lXbrace}}{\llbrace}}
\def\rxbrace{\scalerel*{\usebox{\rXbrace}}{\rrbrace}}
\newcommand{\xbraces}[1]{\scalerel[1.5ex]{\lxbrace}{#1}\scalerel*[1.5ex]{\rxbrace}{#1}}
\title{Implementation and verification of a data structure using dependent types}
\author{Anton Danilkin}
\begin{document}

@@

\begin{no_padding_center}
    \begin{tikzpicture}
        \matrix (node1) [node_matrix] at (0, 0) {
            \coqinlinelight{cons} &
            7 &
            \\
        };
        \matrix (node2) [node_matrix] at (3, 0) {
            \coqinlinelight{cons} &
            5 &
            \\
        };
        \matrix (node3) [node_matrix] at (6, 0) {
            \coqinlinelight{cons} &
            8 &
            \\
        };
        \matrix (node4) [node_matrix] at (9, 0) {
            \coqinlinelight{nil} \\
        };
        \draw [arrow] (node1-1-3.center) to (node2-1-1);
        \draw [arrow] (node2-1-3.center) to (node3-1-1);
        \draw [arrow] (node3-1-3.center) to (node4-1-1);
    \end{tikzpicture}

    \begin{tikzpicture}
        \matrix (node1) [node_matrix] at (0, 0) {
            \coqinlinelight{S} &
            \\
        };
        \matrix (node2) [node_matrix] at (3, 0) {
            \coqinlinelight{S} &
            \\
        };
        \matrix (node3) [node_matrix] at (6, 0) {
            \coqinlinelight{S} &
            \\
        };
        \matrix (node4) [node_matrix] at (9, 0) {
            \coqinlinelight{O} \\
        };
        \draw [arrow] (node1-1-2.center) to (node2-1-1);
        \draw [arrow] (node2-1-2.center) to (node3-1-1);
        \draw [arrow] (node3-1-2.center) to (node4-1-1);
    \end{tikzpicture}
\end{no_padding_center}

@@

@@

\begin{no_padding_center}
    \begin{tikzpicture}
        \matrix (node1) [node_matrix, nodes = {minimum height = 70}] at (0, 0) {
            \node (node1-1-1) [align = left] {\coqinlinelight{Digital} \\ \coqinlinelight{List} \\ \coqinlinelight{Cons}}; &
            \completebinaryleaftreedepthtwo{$0$}{$1$}{$2$}{$3$} &
            \\
        };
        \matrix (node2) [node_matrix, nodes = {minimum height = 40}] at (6, 0) {
            \node (node2-1-1) [align = left] {\coqinlinelight{Digital} \\ \coqinlinelight{List} \\ \coqinlinelight{Cons}}; &
            \O &
            \\
        };
        \matrix (node3) [node_matrix, nodes = {minimum height = 40}] at (10, 0) {
            \node (node3-1-1) [align = left] {\coqinlinelight{Digital} \\ \coqinlinelight{List} \\ \coqinlinelight{Cons}}; &
            \completebinaryleaftreedepthzero{$4$} &
            \\
        };
        \matrix (node4) [node_matrix, nodes = {minimum height = 40}] at (14, 0) {
            \node (node4-1-1) [align = left] {\coqinlinelight{Digital} \\ \coqinlinelight{List} \\ \coqinlinelight{Nil}}; \\
        };
        \draw [arrow] (node1-1-3.center) to (node2-1-1);
        \draw [arrow] (node2-1-3.center) to (node3-1-1);
        \draw [arrow] (node3-1-3.center) to (node4-1-1);
    \end{tikzpicture}
\end{no_padding_center}

@@

\begin{no_padding_center}
    $\xbraces{
        \vcenter{\hbox{\completebinaryleaftreedepthtwo{$0$}{$1$}{$2$}{$3$}}},
        \vcenter{\hbox{\bigvarnothing}},
        \vcenter{\hbox{\completebinaryleaftreedepthzero{$4$}}}
    }$
\end{no_padding_center}

@@

The function \coqinline{to_digits : forall (d n : nat), list nat} gives the list of length \coqinline{d} consisting of digits of \coqinline{i} in big-endian ordering: for example, \coqinline{to_digits 8 73} gives \coqinline{[0; 1; 0; 0; 1; 0; 0; 1]}. The inverse operation is performed by \coqinline{from_digits : forall (d : nat) (nl : list nat), nat}: for example, \coqinline{from_digits 8 [0; 1; 0; 0; 1; 0; 0; 1]} gives \coqinline{73}.

@@

\begin{no_padding_center}
    \begin{tikzpicture}
        \matrix (node1) [node_matrix] at (0, 0) {
            \coqinlinelight{cons} &
            0 &
            \\
        };
        \matrix (node2) [node_matrix] at (3, 0) {
            \coqinlinelight{cons} &
            1 &
            \\
        };
        \matrix (node3) [node_matrix] at (6, 0) {
            \coqinlinelight{cons} &
            2 &
            \\
        };
        \matrix (node4) [node_matrix] at (9, 0) {
            \coqinlinelight{cons} &
            3 &
            \\
        };
        \matrix (node5) [node_matrix] at (12, 0) {
            \coqinlinelight{nil} \\
        };
        \draw [arrow] (node1-1-3.center) to (node2-1-1);
        \draw [arrow] (node2-1-3.center) to (node3-1-1);
        \draw [arrow] (node3-1-3.center) to (node4-1-1);
        \draw [arrow] (node4-1-3.center) to (node5-1-1);
    \end{tikzpicture}
\end{no_padding_center}

>>>>

\documentclass{article}
\usepackage[a4paper, total={19cm, 26cm}]{geometry}
\usepackage[pdfusetitle]{hyperref}
\usepackage{url}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage[backend=biber, style=alphabetic, sorting=none]{biblatex}
\addbibresource{main.bib}
\nocite{*}
\usepackage[usenames, dvipsnames, svgnames, x11names]{xcolor}
\usepackage{graphics}
\usepackage{scalerel}
\usepackage{accsupp}
\usepackage{stmaryrd}
\usepackage{listings}
\usepackage{lstcoq}
\usepackage{tikz}
\usetikzlibrary{positioning}
\usetikzlibrary{matrix}
\usetikzlibrary{bending}
\usetikzlibrary{arrows.meta}
\tikzstyle{node_matrix} = [matrix of nodes, nodes in empty cells, nodes = {anchor = center, minimum size = 20, draw, fill = blue!10, inner sep = 4}, inner sep = 0]
\tikzstyle{arrow} = [[-{>[length = 7, width = 5, bend]}]
\tikzstyle{leaf_tree} = [level distance = 25, level 1/.style = {sibling distance = 40}, level 2/.style = {sibling distance = 20}]
\tikzstyle{leaf_tree_leaf} = [minimum size = 0, circle, draw]
\tikzstyle{leaf_tree_inner_node} = [minimum size = 0, circle, draw, inner sep = 1, fill = black]
\newcommand{\bigvarnothing}[0]{%
    \scalebox{3}{%
        \begin{tikzpicture}%
            \clip (-0.114, -0.1) rectangle (0.114, 0.128);
            \node [style = {inner sep = 0}, text = black!35] {$\varnothing$};
        \end{tikzpicture}%
    }%
}
\newcommand{\completebinaryleaftreedepthzero}[1]{%
    \begin{tikzpicture}%
        \begin{scope} [leaf_tree]
            \node [style = {leaf_tree_leaf}] {#1};
        \end{scope}
    \end{tikzpicture}%
}
\newcommand{\completebinaryleaftreedepthone}[2]{%
    \begin{tikzpicture}%
        \begin{scope} [leaf_tree]
            \node [style = {leaf_tree_inner_node}] {}
                child {
                    node [style = {leaf_tree_leaf}] {#1}
                }
                child {
                    node [style = {leaf_tree_leaf}] {#2}
                };
        \end{scope}
    \end{tikzpicture}%
}
\newcommand{\completebinaryleaftreedepthtwo}[4]{%
    \begin{tikzpicture}%
        \begin{scope} [leaf_tree]
            \node [style = {leaf_tree_inner_node}] {}
                child {
                    node [style = {leaf_tree_inner_node}] {}
                    child {
                        node [style = {leaf_tree_leaf}] {#1}
                    }
                    child {
                        node [style = {leaf_tree_leaf}] {#2}
                    }
                }
                child {
                    node [style = {leaf_tree_inner_node}] {}
                    child {
                        node [style = {leaf_tree_leaf}] {#3}
                    }
                    child {
                        node [style = {leaf_tree_leaf}] {#4}
                    }
                };
        \end{scope}
    \end{tikzpicture}%
}
\newenvironment{no_space_center}
    {\parskip=0pt\par\nopagebreak\centering}
    {\par\noindent\ignorespacesafterend}
\newenvironment{no_padding_center}
    {\parskip=0pt\par\nopagebreak\centering}
    {\par\noindent}
\newcommand*{\llbrace}{%
    \BeginAccSupp{method=hex,unicode,ActualText=2983}%
    \textnormal{\usefont{OMS}{lmr}{m}{n}\char102}%
    \mathchoice{\mkern-4.05mu}{\mkern-4.05mu}{\mkern-4.3mu}{\mkern-4.8mu}%
    \textnormal{\usefont{OMS}{lmr}{m}{n}\char106}%
    \EndAccSupp{}%
}
\newcommand*{\rrbrace}{%
    \BeginAccSupp{method=hex,unicode,ActualText=2984}%
    \textnormal{\usefont{OMS}{lmr}{m}{n}\char106}%
    \mathchoice{\mkern-4.05mu}{\mkern-4.05mu}{\mkern-4.3mu}{\mkern-4.8mu}%
    \textnormal{\usefont{OMS}{lmr}{m}{n}\char103}%
    \EndAccSupp{}%
}
\newsavebox{\lXbrace}
\savebox{\lXbrace}{$\llbrace$}
\newsavebox{\rXbrace}
\savebox{\rXbrace}{$\rrbrace$}
\def\lxbrace{\scalerel*{\usebox{\lXbrace}}{\llbrace}}
\def\rxbrace{\scalerel*{\usebox{\rXbrace}}{\rrbrace}}
\newcommand{\xbraces}[1]{\scalerel[1.5ex]{\lxbrace}{#1}\scalerel*[1.5ex]{\rxbrace}{#1}}
\newcommand{\evaluatesto}[0]{\:\scaleobj{1.5}{\leadsto}\:}
\title{Implementation and verification of a data structure using dependent types}
\author{Anton Danilkin}
\begin{document}

@@

\begin{center}
    \begin{tikzpicture}
        \matrix (node1) [node_matrix, anchor = west] at (0, 0) {
            \coqinlinelight{cons} &
            7 &
            \\
        };
        \matrix (node2) [node_matrix, anchor = west] at (3, 0) {
            \coqinlinelight{cons} &
            5 &
            \\
        };
        \matrix (node3) [node_matrix, anchor = west] at (6, 0) {
            \coqinlinelight{cons} &
            8 &
            \\
        };
        \matrix (node4) [node_matrix, anchor = west] at (9, 0) {
            \coqinlinelight{nil} \\
        };
        \draw [arrow] (node1-1-3.center) to (node2-1-1);
        \draw [arrow] (node2-1-3.center) to (node3-1-1);
        \draw [arrow] (node3-1-3.center) to (node4-1-1);
    \end{tikzpicture}
\end{center}

\begin{center}
    \begin{tikzpicture}
        \matrix (node1) [node_matrix, anchor = west] at (0, 0) {
            \coqinlinelight{S} &
            \\
        };
        \matrix (node2) [node_matrix, anchor = west] at (3, 0) {
            \coqinlinelight{S} &
            \\
        };
        \matrix (node3) [node_matrix, anchor = west] at (6, 0) {
            \coqinlinelight{S} &
            \\
        };
        \matrix (node4) [node_matrix, anchor = west] at (9, 0) {
            \coqinlinelight{O} \\
        };
        \draw [arrow] (node1-1-2.center) to (node2-1-1);
        \draw [arrow] (node2-1-2.center) to (node3-1-1);
        \draw [arrow] (node3-1-2.center) to (node4-1-1);
    \end{tikzpicture}
\end{center}

@@

\begin{center}
    \begin{tikzpicture}
        \matrix (node1) [node_matrix, nodes = {minimum height = 40}, anchor = center] {
            \node (node1-1-1) [align = left] {\coqinlinelight{Binary} \\ \coqinlinelight{LeafTree} \\ \coqinlinelight{InternalNode}}; &
            &
            \\
        };
        \matrix (node2) [node_matrix, nodes = {minimum height = 40}, anchor = center, node distance = 1.5 and 1.5, below left = of node1-1-2] {
            \node (node2-1-1) [align = left] {\coqinlinelight{Binary} \\ \coqinlinelight{LeafTree} \\ \coqinlinelight{InternalNode}}; &
            &
            \\
        };
        \matrix (node3) [node_matrix, nodes = {minimum height = 40}, anchor = center, node distance = 1.5 and 1.5, below right = of node1-1-3] {
            \node (node3-1-1) [align = left] {\coqinlinelight{Binary} \\ \coqinlinelight{LeafTree} \\ \coqinlinelight{InternalNode}}; &
            &
            \\
        };
        \matrix (node4) [node_matrix, nodes = {minimum height = 40}, anchor = center, node distance = 1.5 and 0.1, below left = of node2-1-2] {
            \node (node4-1-1) [align = left] {\coqinlinelight{Binary} \\ \coqinlinelight{LeafTree} \\ \coqinlinelight{Leaf}}; &
            $0$ \\
        };
        \matrix (node5) [node_matrix, nodes = {minimum height = 40}, anchor = center, node distance = 1.5 and 0.1, below right = of node2-1-3] {
            \node (node5-1-1) [align = left] {\coqinlinelight{Binary} \\ \coqinlinelight{LeafTree} \\ \coqinlinelight{Leaf}}; &
            $1$ \\
        };
        \matrix (node6) [node_matrix, nodes = {minimum height = 40}, anchor = center, node distance = 1.5 and 0.1, below left = of node3-1-2] {
            \node (node6-1-1) [align = left] {\coqinlinelight{Binary} \\ \coqinlinelight{LeafTree} \\ \coqinlinelight{Leaf}}; &
            $2$ \\
        };
        \matrix (node7) [node_matrix, nodes = {minimum height = 40}, anchor = center, node distance = 1.5 and 0.1, below right = of node3-1-3] {
            \node (node7-1-1) [align = left] {\coqinlinelight{Binary} \\ \coqinlinelight{LeafTree} \\ \coqinlinelight{Leaf}}; &
            $3$ \\
        };
        \draw [arrow] (node1-1-2.center) -- (node1-1-2.south) to[out = -90, in = 90] (node2.north);
        \draw [arrow] (node1-1-3.center) -- (node1-1-3.south) to[out = -90, in = 90] (node3.north);
        \draw [arrow] (node2-1-2.center) -- (node2-1-2.south) to[out = -90, in = 90] (node4.north);
        \draw [arrow] (node2-1-3.center) -- (node2-1-3.south) to[out = -90, in = 90] (node5.north);
        \draw [arrow] (node3-1-2.center) -- (node3-1-2.south) to[out = -90, in = 90] (node6.north);
        \draw [arrow] (node3-1-3.center) -- (node3-1-3.south) to[out = -90, in = 90] (node7.north);
    \end{tikzpicture}
\end{center}

They will be drawn like that in the future:

\begin{center}
    \completebinaryleaftreedepthtwo{$0$}{$1$}{$2$}{$3$}
\end{center}

@@

\begin{center}
    \begin{tikzpicture}
        \matrix (node1) [node_matrix, nodes = {minimum height = 70}, anchor = west] at (0, 0) {
            \node (node1-1-1) [align = left] {\coqinlinelight{Digital} \\ \coqinlinelight{List} \\ \coqinlinelight{Cons}}; &
            \completebinaryleaftreedepthtwo{$0$}{$1$}{$2$}{$3$} &
            \\
        };
        \matrix (node2) [node_matrix, nodes = {minimum height = 40}, anchor = west] at (6, 0) {
            \node (node2-1-1) [align = left] {\coqinlinelight{Digital} \\ \coqinlinelight{List} \\ \coqinlinelight{Cons}}; &
            \O &
            \\
        };
        \matrix (node3) [node_matrix, nodes = {minimum height = 40}, anchor = west] at (10, 0) {
            \node (node3-1-1) [align = left] {\coqinlinelight{Digital} \\ \coqinlinelight{List} \\ \coqinlinelight{Cons}}; &
            \completebinaryleaftreedepthzero{$4$} &
            \\
        };
        \matrix (node4) [node_matrix, nodes = {minimum height = 40}, anchor = west] at (14, 0) {
            \node (node4-1-1) [align = left] {\coqinlinelight{Digital} \\ \coqinlinelight{List} \\ \coqinlinelight{Nil}}; \\
        };
        \draw [arrow] (node1-1-3.center) to (node2-1-1);
        \draw [arrow] (node2-1-3.center) to (node3-1-1);
        \draw [arrow] (node3-1-3.center) to (node4-1-1);
    \end{tikzpicture}
\end{center}

@@

\begin{center}
    $\xbraces{
        \vcenter{\hbox{\completebinaryleaftreedepthtwo{$0$}{$1$}{$2$}{$3$}}};
        \vcenter{\hbox{\bigvarnothing}};
        \vcenter{\hbox{\completebinaryleaftreedepthzero{$4$}}}
    }$
\end{center}

@@

The function \coqinline{to_digits : forall (d n : nat), list nat} gives the list of length \coqinline{d} consisting of digits of \coqinline{i} in big-endian ordering: for example, \coqinline{to_digits 8 73 ~> [0; 1; 0; 0; 1; 0; 0; 1]}. The inverse operation is performed by \coqinline{from_digits : forall (d : nat) (nl : list nat), nat}: for example, \coqinline{from_digits 8 [0; 1; 0; 0; 1; 0; 0; 1] ~> 73}.

@@

\begin{no_padding_center}
    $[0; 1; 2; 3]$
\end{no_padding_center}
